---
title: "Basic R"
editor: visual
---

# Component overview

[R](https://cran.r-project.org/) is a free and open source statistical computing language. [RStudio](https://posit.co/products/open-source/rstudio/) is an integrated development environment (IDE) that makes using R much easier. You need to install both on your computer to get started.

# Installing R, RStudio (UWSC)

-   Open Software Center
-   Click to install R
-   Click to install RStudio
-   Install packages

# Installing R, RStudio (Non-UWSC)

-   Install latest R from [R's main site](https://cran.r-project.org/)

-   Install RStudio from [Posit's site](https://posit.co/downloads/)

-   Install packages

# Orientation to the interface

-   Square: Code, Console, Environment, Files
-   Special views
-   Pop out code

# Workflow

Example getting new data from Qualtrics/POD, process for delivery, zip, post to Box

```{r, eval = FALSE}

df <- read_spss("./raw/PXXX Qualtrics Data File Raw.sav")
df <- as_factor(df)

# Clean, etc
df <- df %>% 
  select(
    csid = ExternalReference,
    Q1:Q35
  )


write_csv(df, "./out/PXXX Qualtrics Data File Clean.csv")
write_sav(df, "./out/PXXX Qualtrics Data File Clean.sav")
write_dta(df, "./out/PXXX Qualtrics Data File Clean.dta")

zip(
  paste0("./pst/PXXX Data delivery for ", Sys.Date(), ".zip"),
  c(
    "./out/PXXX Qualtrics Data File Clean.csv",
    "./out/PXXX Qualtrics Data File Clean.sav",
    "./out/PXXX Qualtrics Data File Clean.dta"
  )
)

box_auth()

box_ul(
  dir_id = 128932912815,
  paste0("./pst/PXXX Data delivery for ", Sys.Date(), ".zip"),
)

```

# Helper packages: tidyverse, haven, boxr, qualtRics

-   Tidyverse - Lots of helpful packages like dplyr, tidyr, stringr, etc
-   haven - Reads and writes SPSS, Stata, csv
-   boxr - Posts to Box
-   qualtRics - Reads from Qualtrics

# Importing/exporting various formats

```{r, eval=FALSE}
library(haven)

read_spss("./raw/Any SPSS File.sav")
read_dta("./raw/Any Stata File.dta")
read_csv("./raw/Any csv File.csv")

library(readxl)

read_xlsx("./raw/Any Excel File Skip First Two Rows.xlsx", skip=2)

```
