---
title: "Basic R"
editor: visual
---

# Why use R?

R and RStudio have several advantages over statistical software like Stata, SAS, and SPSS.

-   R can load multiple data sources in multiple formats at the same time.
-   R is open source and helper packages are widely developed by users.
-   R incorporates version control and collaboration tools like git and GitHub.
-   R can output data to any file format and can create reports in HTML, PDF, docx, and pptx format.
-   The RStudio IDE has a nice organization tool called [Projects](https://nrjjones.github.io/projects.html).

# Component overview

[R](https://cran.r-project.org/) is a free and open source statistical computing language. [RStudio](https://posit.co/products/open-source/rstudio/) is an integrated development environment (IDE) that makes using R much easier. You need to install both on your computer to get started.

# Installing R, RStudio (UWSC)

-   Open Software Center
-   Click to install R
-   Click to install RStudio
-   (Possibly necessary) Open RGui to initialize local folders
-   Install packages

# Installing R, RStudio (Non-UWSC)

-   Install latest R from [R's main site](https://cran.r-project.org/)
-   Install RStudio from [Posit's site](https://posit.co/downloads/)
-   Install packages

# Orientation to the interface

-   Grid: Code, Console, Environment, Files
-   Special views
-   Pop out code

# Helper packages:

-   tidyverse - Lots of helpful data manipulation tools
-   haven - Reads and writes SPSS, Stata, csv
-   boxr - Posts files to Box
-   qualtRics - Reads and processes data from Qualtrics

# Importing/exporting various formats

```{r, eval=FALSE}

library(haven)

df1 <- read_spss("./raw/Any SPSS File.sav")
df2 <- read_dta("./raw/Any Stata File.dta")
df3 <- read_csv("./raw/Any csv File.csv")

library(readxl)

df4 <- read_xlsx("./raw/Any Excel File Skip First Two Rows.xlsx", skip=2)

```

# Example workflow

One way of getting new data from Qualtrics/POD, process for delivery, zip, post to Box

```{r, eval = FALSE}

# Read survey data file and convert to labeled data

df <- read_spss("./raw/PXXX Wave 1 Data File Raw.sav")
df <- as_factor(df)

# Read sample file and keep important sample data

sam <- read_csv("./raw/PXXX Sample File.csv")

sam <- sam %>% 
  filter(wave == 1) %>% 
  select(csid, dob, zip)

# Clean, remove extra variables, etc

df <- df %>% 
  filter(!is.na(csid)) %>% 
  select(
    csid = ExternalReference,
    Q1:Q35
  )

# Merge survey with sample data

df <- left_join(df, sam, by="csid")

# Write data in csv, SPSS, and Stata formats

write_csv(df, "./out/PXXX Interim Data w Sample Info.csv")
write_sav(df, "./out/PXXX Interim Data w Sample Info.sav")
write_dta(df, "./out/PXXX Interim Data w Sample Info.dta")

# Zip into folder with date stamp

zip(
  paste0("./pst/PXXX Interim Data delivery for ", Sys.Date(), ".zip"),
  c(
    "./out/PXXX Interim Data w Sample Info.csv",
    "./out/PXXX Interim Data w Sample Info.sav",
    "./out/PXXX Interim Data w Sample Info.dta"
  )
)

# Refresh Box athenitication

box_auth()

# Upload date stamp to zip

box_ul(
  dir_id = 1234567891234,
  paste0("./pst/PXXX Interim Data delivery for ", Sys.Date(), ".zip"),
)

```
