---
title: "Sample Cleaning"
editor: visual
---

# Check for duplicates

```{r, eval=FALSE}

tmp %>% df %>% group_by(email) %>% mutate(n=n()) %>% filter(n>1)
```

# Add csid

```{r, eval=FALSE}
df <- df %>% 
  mutate(
    csid = 1000 + row_number()
  )
```

# Uncap names

```{r, eval=FALSE}
df <- df %>% 
  mutate(
    first.orig = first,
    first = str_to_title(first)
  )
```

# Fix secondary problems from uncapping names

# Separate names from single field

```{r, eval=FALSE}
df <- df %>% 
  mutate(
    first = str_extract(fullname, ", [A-Za-z]*"),
    first = str_remove(first, ", "),
    first = str_extract(fullname, "[A-Za-z], "),
    last = str_remove(last, ", ")
  )
```

# PENDING

# Separate address parts from a single field

# Fix other common problems:

## McName

## O'Name

## Ii, Iii, jr, etc

## Po Box, Us Hwy

## Directionals: Se/Nw

## 1st/2nd

## Rural roads: N123e1234

## Of, Du, De

# Format output

# Sample in a reproducible way using set.seed()

# Post to Box

Related: OCR to scan text from PDF, extract information from a formatted document (eg Amazon codes, address, etc)\
