---
title: " "
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: 
        collapsed: false
---

# Stata Rosetta Stone

It takes time to start thinking like an R programmer instead of a Stata programmer.  This section shows how to do some common Stata tasks using R.  

# Working with data in general

Unlike Stata, you can have multiple datasets open at the same time in R.  Unfortunately, this flexibily requires you to state *explicitly* where the variables are every time you use them.  For example, if you have a sample file with personal information and another data file with survey responses, you might load both data files with the names **sample** and **survey**.  If you have linking variable in both called **ID**, you refer to them with the data frame name then a dollar sign (\$) and then the variable like this:   **sample\$ID** or **survey\$ID**.    

In addition to their names, you can access variables by their column number in brackets.  For example, if ID is the first column in both data files, you could type **sample[1]** or **survey[1]** instead of the **\$ID** format used above.  You can also use quoted variable names in brackets if the variable name has spaces, for example, **sample['street address line 1']**.

If you are only using one data file for your project, most of the R world finds it convenient to use the abbreviation **df**, short for *dataframe*, for the main data file.  If you need to load multiple data files, use the shortest reference name you can to reduce your typing burden.  For example, you can save some extra typing by using **sur** instead of **survey** and **sam** instead of **sample**.  

# Recoding data

## Simple recoding and IF statements

Stata code requires conditional statements after the action.  For example: 

>   replace state = "GA" if state == "Georgia"

R has a convention of putting the condition in brackets attached to the dataframe and variable name.  The following code can be interpreted as 'For the variable **state** in the dataframe **df**, when the variable **state** in **df** equals "Georgia", replace the value with "GA".   

```{r, eval=FALSE}
   df$state[df$state=="Georgia"] <- "GA"  
```

An example with more more than one condition:  

```{r, eval=FALSE}
   df$salutation[df$md_degree=="Yes" | df$educ=="PhD"] <- "Dr. "  
```

## Complex recoding data

If you need to do more complex recoding, use case_when() from dplyr:   

```{r, eval=FALSE}
df$state <- case_when(  
    df$state == "Georgia" ~ "GA",  
    df$state == "Florida" ~ "FL",  
    df$state == "Wisconsin" ~ "WI",  
    df$state == "Wyoming" ~ "WY",  
    )
```
   
# Loops

## Vectorized apply vs for loop

Stata users are very familiar with a range of loop functions (foreach, forval, etc).  You can write loops in R, but some problems that loops solve in Stata are better handled by vectorized code in R.  

Some resources: 

- [A Tutorial on Loops in R](https://www.datacamp.com/community/tutorials/tutorial-on-loops-in-r)  
- [Vectorization in R: Why?](http://www.noamross.net/blog/2014/4/16/vectorization-in-r--why.html)  
- [Let's talk about vectorization](http://alyssafrazee.com/2014/01/29/vectorization.html)  
- [Loops in R](https://swcarpentry.github.io/r-novice-inflammation/15-supp-loops-in-depth/)  

Placeholder for get() to approximate for loop functionality
   
# Merging data

* **full_join() ** - similar to merge, but no _merge variable is created.  Consider installing **tidylog** or running your own checks to produce a variable similar _merge  

* **left_join() ** - similar to merge, then keep the _merge==3 

* **anti_join() ** - similar to merge, then keep the _merge==1  | _merge==2

## Reshaping Data 

Many of R's popular graphing and mapping tools prefer data to be in long format.  Hadley Wickham has [written eloquently](https://vita.had.co.nz/papers/tidy-data.pdf) about why long data, or *tidy* data, has many advantages. If you frequently work with social science data files that tend to be stored in wide files, Knowing how to reshape your data is an essential skill.

### Wide to long

See [here](https://stats.idre.ucla.edu/stata/modules/reshaping-data-wide-to-long/) for a good example of reshaping wide data to long format.  

Wide format
```{r, eval=FALSE}
id      inc96      inc97      inc98   
1003    75000      76000      77000    
1004    40000      40500      41000    
1005    45000      45400      45800    
```

Long format
```{r, eval=FALSE}
id        year    inc 
1003      96      40000  
1003      97      40500  
1003      98      41000  
1004      96      45000  
1004      97      45400  
1004      98      45800  
1005      96      75000  
1005      97      76000  
1005      98      77000 
```

Stata's reshape command

Stata parses the variable name to separate inc97 into the year and the income value.  

```{r, eval=FALSE}
reshape long inc, i(id) j(year) 
```

gather() from the **tidyr** package works very similarly, but leaves the value of year as inc97, etc.  You can change these to a year value after.  

```{r, eval=FALSE}
df_long <- gather(df, year, inc, -id)

# To leave 96, but as a character variable
df_long <- str_remove(df_long$year, "inc")

# To leave 96 as a number
df_long <- as.numeric(str_remove(df_long$year, "inc")) 

# To leave 1996 as a number
df_long <- as.numeric(str_remove(df_long$year, "inc"))+1900 
```

### Long to wide

Stata

```{r, eval=FALSE}
reshape wide age, i(id)  j(year) 
```

spread() from the tidyr package

```{r, eval=FALSE}
df_wide <- spread(df_long, year, inc)

# But this leaves the column names as 96, 97, 98. To rename:
names(df_wide) <- c("id", "inc96", "inc97", "inc98")
```


